---
title: "R Downloads"
author: "A Gruer"
date: "30 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and theme, echo=TRUE}
library(tidyverse)
library(lubridate)
library(qicharts2)
library(countrycode)
theme_set(theme_minimal())
```


```{r download data, eval=FALSE, include=FALSE}
dir.create("data")
download.file("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-30/r_downloads_year.csv", "data/r_downloads_year.csv")
```
```{r read data }
r_downloads_raw <- read.csv("data/r_downloads_year.csv", stringsAsFactors = FALSE)

r_downloads <- r_downloads_raw %>% 
                    select(-X) %>%
                    mutate(time = as.POSIXct(paste(date,time),tz="UTC"),
                          date = as.Date(date)
                           )

```

There were `r nrow(r_downloads_raw) %>% scales::comma()` downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018.

Which versions of R were downloaded 
Why is 2.1 the most downloaded 2 version?
```{r downloads by version}
r_downloads %>%
  mutate(major_version = substr(version,1,3)) %>%
  mutate(major_version =  fct_rev( fct_infreq(major_version))) %>%
   mutate(major_version =  fct_recode(major_version,latest = "lat", development = "dev" )) %>% 
  count(major_version) %>%
  filter(n > 1000) %>%
 
  ggplot(aes(major_version, y = n, fill = major_version)) +
  geom_col(  alpha = 0.75) +
  scale_y_log10(labels=scales::comma) +
  #theme(axis.text.x = element_text(angle=45,hjust=1)) +
  theme(panel.grid.major.y =  element_blank()) +
  coord_flip()+
  labs(title = "Annual downloads of R by major version (more than 1,000 downloads)",
       subtitle = "log 10 scale",
       caption = "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018")
  
```

Which Operating Systems?
Windows is by far the most represented operating system

```{r operating system}
r_downloads %>%
  mutate(os =  fct_infreq(os)) %>%
  ggplot(aes(os, fill = os)) +
  geom_bar(show.legend = FALSE, alpha = 0.75) +
  scale_y_continuous(labels=scales::comma) +
  theme(panel.grid = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_text(size = rel(2))) +
  labs(title = "Annual downloads of R by OS",
       caption =  "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018" ) 
  
```

Time series of downloads per day

```{r}
r_downloads %>%
  
  count(date) %>%
  ggplot(aes(x = date, y = n )) + 
  geom_line()
```

Time series of downloads per day by version

```{r time series by version}
major_versions <- 
  r_downloads %>%
  mutate(major_version = substr(version,1,3)) %>%
  mutate(major_version =  fct_rev( fct_infreq(major_version))) %>%
  count(major_version) %>%
  filter(n > 1000) %>%
  pull(major_version)

  r_downloads %>%
  mutate(major_version = substr(version,1,3)) %>%
  mutate(major_version =    fct_infreq(major_version))  %>%
  filter(major_version %in% major_versions) %>%
    mutate(major_version =  fct_recode(major_version,latest = "lat", development = "dev" )) %>% 
  count(date,major_version) %>%
  ggplot(aes(x = date, y = n )) + 
  geom_line() +
 
      theme(axis.text.x = element_text(angle=45,hjust=1),
        panel.grid = element_blank(),
        strip.text.x = element_text(face = "bold",size=rel(1.5))
        
        )+
    scale_y_continuous(labels=scales::comma)+
  facet_wrap(~major_version, scale="free_y") +
    labs(title = "R Downloads per day - 9 most commonly downloaded major versions",
 
          
       caption =  "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018") 
  
  
```

By Country

```{r downloads by country}
 
r_downloads %>%
  mutate(country = fct_infreq(country)) %>%
  count(country) %>%
  head(9) %>%
  mutate(countryname = countrycode(country,"iso2c","country.name") )  %>%
  mutate(countryname = fct_reorder(countryname, n)) %>%
  ggplot(aes(countryname, n, fill = countryname)) +
  geom_col(show.legend = FALSE, alpha = 0.75) +
  scale_y_continuous(labels=scales::comma) +
  coord_flip()+
  theme(panel.grid = element_blank(),
        panel.grid.minor.x = element_blank()) +
   labs(title = "Annual downloads of R by Country",
       caption =  "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018", x = "" ) 
```

Download by continent

```{r downloads by continent}
 
r_downloads %>%
  mutate(continent =       countrycode(country,"iso2c","continent") )  %>%
    mutate(country_name =       countrycode(country,"iso2c","country.name") )  %>%
  count(continent) %>%
  mutate(continent = fct_reorder(continent, n)) %>%
  ggplot(aes(continent, n, fill = continent)) +
  geom_col(show.legend = FALSE, alpha = 0.75) +
  scale_y_continuous(labels=scales::comma) +
  coord_flip()+
  theme(panel.grid = element_blank(),
        panel.grid.minor.x = element_blank()) +
   labs(title = "Annual downloads of R by Continent",
       caption =  "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018", x = "" )

```

 
 
```{r top 9 countries}
top_9_countries <-
      r_downloads %>%
        mutate(country = fct_infreq(country)) %>%
        count(country) %>%
        head(9) %>%
        pull(country)
        
   r_downloads %>%
   filter(country %in% top_9_countries) %>%
   mutate(country = fct_infreq(country)) %>%
  mutate(countryname = countrycode(country,"iso2c","country.name") )  %>%
  mutate(countryname = fct_infreq(countryname)) %>%
   count(countryname, date) %>%

  ggplot(aes(date, n)) +
  geom_line( )  +
 
      theme(axis.text.x = element_text(angle=45,hjust=1),
        panel.grid = element_blank(),
        strip.text.x = element_text(face = "bold",size=rel(1.5))
        
        )+
    scale_y_continuous(labels=scales::comma)+
  facet_wrap(~countryname, scale="free_y") +
    labs(title = "R Downloads per day - 9 countries with the most annual downloads",
 
          
       caption =  "source: downloads of R from the RStudio CRAN mirror between October 20, 2017 and October 20, 2018") 
  

```
 
